cardinalR: Generating interesting high-dimensional data structures

### 2025-06-24 paper-cardinalR.Rmd
Updated cube section

Each variable $X_j$ represents integer grid coordinates (e.g., 1, 2, 3, ...) in dimension $j$:

$$
X_j \in {1, \dots, k_j},\text{ where }\prod_{j=1}^p k_j \approx n.
$$

This results in a hypercube-shaped lattice in $p$-dimensional space.

@@

The `gen_unifcube(n, p)` function generates $n$ points uniformly distributed within a $3\text{-}D$ cube centered at the origin, optionally embedded in a higher-dimensional space with additional noise dimensions.

Points are sampled independently from $X_1, X_2, X_3 \sim U(-0.5, 0.5)$ of a cube of side length 1. This results in a uniform cloud of points filling the unit cube in $3\text{-}D$ space.

For $p > 3$, the remaining dimensions are filled with Gaussian noise:

$$
X_j \sim N(0, 0.05^2), \quad j = 4, \dots, p.
$$

@@

The `gen_cubehole(n, p)` function generates a dataset of $n$ points sampled uniformly inside a cube, then removes points that fall within a central spherical hole, resulting in a hollow cube structure in $p$-dimensional space.

The initial points are drawn from a uniform cube centered at the origin: $X_1, X_2, X_3 \sim U(-0.5, 0.5)$.

The center of the cube is $(0, 0, 0)$. A spherical hole of radius $0.5$ is carved out by filtering out points whose Euclidean distance from the origin is less than or equal to $0.5$:

Remove if $\sqrt{X_1^2 + X_2^2 + X_3^2} \leq 0.5$.

This results in a shell-like point cloud that fills the cube but avoids the spherical center.

For $p > 3$, Gaussian noise is added to the remaining dimensions (via `gen_unifcube()`):

$$
X_j \sim N(0, 0.05^2), \quad j = 4, \dots, p.
$$.

>>>>

Each variable $X_j$ represents integer grid coordinates (e.g., 1, 2, 3, ...) in dimension $j$: $X_j \in {1, \dots, k_j},\text{ where }\prod_{j=1}^p k_j \approx n$. This results in a hypercube-shaped lattice in $p\text{-}D$.

@@

The `gen_unifcube(n, p)` function generates $n$ points uniformly distributed within a $3\text{-}D$ cube centered at the origin, optionally embedded in $p\text{-}D$ with additional noise dimensions.

Points are sampled independently from $X_1, X_2, X_3 \sim U(-0.5, 0.5)$ of a cube of side length $1$. This results in a uniform cloud of points filling the unit cube in $3\text{-}D$ space. For $p > 3$, Gaussian noise $X_j \sim N(0, 0.05^2)$ is added to embed the $3\text{-}D$ structure into $p\text{-}D$, where $j = 4, \dots, p$.

@@

The `gen_cubehole(n, p)` function generates a dataset of $n$ points sampled uniformly inside a cube, then removes points that fall within a central spherical hole, resulting in a hollow cube structure in $p\text{-}D$.

The initial points are drawn from a uniform cube centered at the origin: $X_1, X_2, X_3 \sim U(-0.5, 0.5)$. The center of the cube is $(0, 0, 0)$. A spherical hole of radius $0.5$ is carved out by filtering out points whose Euclidean distance from the origin is less than or equal to $0.5$: Remove if $\sqrt{X_1^2 + X_2^2 + X_3^2} \leq 0.5$. This results in a shell-like point cloud that fills the cube but avoids the spherical center. For $p > 3$, Gaussian noise $X_j \sim N(0, 0.05^2)$ is added to embed the $3\text{-}D$ structure into $p\text{-}D$, where $j = 4, \dots, p$.

### 2025-06-24 paper-cardinalR.tex
Updated cube section

Each variable \(X_j\) represents integer grid coordinates (e.g., 1, 2, 3, \ldots) in dimension \(j\):

\[
X_j \in {1, \dots, k_j},\text{ where }\prod_{j=1}^p k_j \approx n.
\]

This results in a hypercube-shaped lattice in \(p\)-dimensional space.

@@

The \texttt{gen\_unifcube(n,\ p)} function generates \(n\) points uniformly distributed within a \(3\text{-}D\) cube centered at the origin, optionally embedded in a higher-dimensional space with additional noise dimensions.

Points are sampled independently from \(X_1, X_2, X_3 \sim U(-0.5, 0.5)\) of a cube of side length 1. This results in a uniform cloud of points filling the unit cube in \(3\text{-}D\) space.

For \(p > 3\), the remaining dimensions are filled with Gaussian noise:

\[
X_j \sim N(0, 0.05^2), \quad j = 4, \dots, p.
\]

@@

The \texttt{gen\_cubehole(n,\ p)} function generates a dataset of \(n\) points sampled uniformly inside a cube, then removes points that fall within a central spherical hole, resulting in a hollow cube structure in \(p\)-dimensional space.

The initial points are drawn from a uniform cube centered at the origin: \(X_1, X_2, X_3 \sim U(-0.5, 0.5)\).

The center of the cube is \((0, 0, 0)\). A spherical hole of radius \(0.5\) is carved out by filtering out points whose Euclidean distance from the origin is less than or equal to \(0.5\):

Remove if \(\sqrt{X_1^2 + X_2^2 + X_3^2} \leq 0.5\).

This results in a shell-like point cloud that fills the cube but avoids the spherical center.

For \(p > 3\), Gaussian noise is added to the remaining dimensions (via \texttt{gen\_unifcube()}):

\[
X_j \sim N(0, 0.05^2), \quad j = 4, \dots, p.
\].

>>>>

Each variable \(X_j\) represents integer grid coordinates (e.g., 1, 2, 3, \ldots) in dimension \(j\): \(X_j \in {1, \dots, k_j},\text{ where }\prod_{j=1}^p k_j \approx n\). This results in a hypercube-shaped lattice in \(p\text{-}D\).

@@

The \texttt{gen\_unifcube(n,\ p)} function generates \(n\) points uniformly distributed within a \(3\text{-}D\) cube centered at the origin, optionally embedded in \(p\text{-}D\) with additional noise dimensions.

Points are sampled independently from \(X_1, X_2, X_3 \sim U(-0.5, 0.5)\) of a cube of side length \(1\). This results in a uniform cloud of points filling the unit cube in \(3\text{-}D\) space. For \(p > 3\), Gaussian noise \(X_j \sim N(0, 0.05^2)\) is added to embed the \(3\text{-}D\) structure into \(p\text{-}D\), where \(j = 4, \dots, p\).

@@

The \texttt{gen\_cubehole(n,\ p)} function generates a dataset of \(n\) points sampled uniformly inside a cube, then removes points that fall within a central spherical hole, resulting in a hollow cube structure in \(p\text{-}D\).

The initial points are drawn from a uniform cube centered at the origin: \(X_1, X_2, X_3 \sim U(-0.5, 0.5)\). The center of the cube is \((0, 0, 0)\). A spherical hole of radius \(0.5\) is carved out by filtering out points whose Euclidean distance from the origin is less than or equal to \(0.5\): Remove if \(\sqrt{X_1^2 + X_2^2 + X_3^2} \leq 0.5\). This results in a shell-like point cloud that fills the cube but avoids the spherical center. For \(p > 3\), Gaussian noise \(X_j \sim N(0, 0.05^2)\) is added to embed the \(3\text{-}D\) structure into \(p\text{-}D\), where \(j = 4, \dots, p\).
